---
title: "Teffer Hot Sick Trout Project"
author: "Amy Teffer"
date: "5/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, echo = TRUE)
```

This Smith Project aims to investigate the interacting impacts of elevated water temperature, pathogen dynamics, and increased connectivity on brook trout resilience into the future.

This modeling exercise is meant to:
* build my skills in matrix population modeling
* identify parameters and variables that will have a large impact on population size over time and relative magnitude of effects across a range of values

# Practicing Matrix Population Modeling

```{r, include=T, results="hide", message=F}
#Load packaages and set directory
library(expm)    
library(ggplot2)
library(dplyr)
library(reshape2)
#setwd("~/Smith Conservation Fellow/Smith BKT Data/R Smith")

## Create our matrix
mtx <- matrix(
  c(0.0,1.0,0.0,0.4,
    0.4,0.3,0.1,0.3,
    0.0,0.6,0.0,0.9,
    0.2,0.3,0.4,0.5),
  nrow=4)
mtx

## Create vector of age distributions, Pops A&B
pop.age<-c(20,50,40,30)
pop.age

## Multiply matrix by vector for one cycle
result.1cyc <- mtx %*% pop.age
result.1cyc

result.2cyc <- (mtx %^% 2) %*% pop.age
result.2cyc
```

## Loop by year and store values in matrix

```{r}
### Create empty matrix to store values per year for 100 years
mat <- matrix(,nrow=4, ncol=100)

### Loop over 100 years
for(i in 1:100){
      mat[,i] = (mtx %^% i) %*% pop.age
  }

### Plot age distributions in 2 populations over time
pop.100 <- as.data.frame(mat)
rownames(pop.100) <- c("PopA0", "PopA1", "PopB0", "PopB1")
colnames(pop.100) <- c(1:100)

### Add row names and melt into long form to plot
pop.df <- tibble::rownames_to_column(pop.100, "pop")
pop.df2 <- melt(data = pop.df,
                id.vars = "pop")

### Plot population size change over 100 cycles
ggplot(pop.df2, aes(x=variable, y=value, fill=pop)) + 
  geom_bar(stat = "identity") +
  labs(y="Population size", x="Cycle") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Fig.1 Stacked bar chart of population sizes over time for 2 populations (A,B) and 2 age classes each (0, 1+). Why is my population exploding into space??
